\documentclass[11pt]{exam}
\usepackage[empty]{fullpage}
\usepackage{amsmath,amssymb}
\usepackage{tikz}
\usepackage{color,comment,soul}

\newcommand{\blank}[1]{\underline{\hspace*{#1}}}
\newcommand{\ds}{\displaystyle}
\newcommand{\R}{\mathbb{R}}
\newcommand{\inner}[1]{\langle {#1} \rangle}



\begin{document}


\subsection*{Numerical Analysis - Math 342 \hfill Final Exam Review}

\textit{The final exam will be on \textbf{Thursday, April 30 at 9:00am}. The following problems are similar to ones you might see on the final exam.}


\begin{questions}
\question Let $f\left(x\right)\ =\ x^{3\ }+2x-8$.    
\begin{parts}
\part Apply Newton's method twice to $f(x)$ starting with $x_0 = 1$. What are $x_1$ and $x_2$? 
\begin{solution}
$$x_1 = 2 \text{ and } x_2 = \tfrac{12}{7}.$$
\end{solution}
\vfill

\part If you were to continue applying Newton's method in this case, the sequence $x_k$ would converge to a number $x_\infty$. What is $f(x_\infty)$? 
\begin{solution}
The limit of Newton's method (if it converges) is a root of $f$, so $f(x_\infty) = 0$.  
\end{solution}
\vfill

\part If you start with $x_0 = 0$ and $x_1 = 1$, then what is $x_2$ according to the secant method?
\begin{solution}

\end{solution}
\vfill
\end{parts}



\question Consider the four points $(-2,0), (0, 6), (1,18), (3,30)$.

\begin{parts}
\part Write the interpolating polynomial using Lagrange polynomials. 
\begin{solution}
\end{solution}
\vfill

\part Make a table of divided differences for these four points.
\begin{solution}
\end{solution}
\vfill
\vfill

\part Write the interpolating polynomial using the Newton basis.
\vfill

\end{parts}

\newpage

\question Use the Simpson's method error formula 
$$|\text{Error}| \le \max_{a \le \xi \le b} \frac{|f^{(4)}(\xi)(b-a)^5}{2880 n^4}$$
to find the minimum number of subintervals needed to guarantee that the you can approximate the integral $\int_0^2 e^{-x^2} \, dx$ with an absolute error less than $10^{-8}$. Hint: if $f(x) = e^{-x^2}$, then the maximum value of $|f^{(4)}(\xi)|$ is 12 on the interval $[0,2]$.  


\begin{solution}
\end{solution}
\vfill

\question Use Euler's method to approximate the solution of the initial value problem 
$$\ds \frac{dy}{dx} = x - y^2 \text{ on } [0,3] \text{ with initial condition } y(0) = 1.$$  
Use $n=3$ and fill in the table below.

\begin{center}
\renewcommand{\arraystretch}{2.0}
\begin{tabular}{|c|c|c|c|c|}
\hline
~~~~$x$~~~~ & 0 &  \hspace*{0.8in} & \hspace*{0.8in} & \hspace*{0.8in} \\ \hline
$y$ & \hspace*{0.8in}  & & & \\ \hline
\end{tabular}
\end{center}

\begin{solution}
\end{solution}
\vfill

\question Compute the LU-decomposition of $A = \begin{pmatrix} 1 & 0 & 0 \\ 1 & -1 & 1 \\ 1 & 2 & 4 \end{pmatrix}$. 
\begin{solution}
$$L = \begin{pmatrix} 1 & 0 & 0 \\ 1 & 1 & 0 \\ 1 & -2 & 1\end{pmatrix}, ~~~~ U = \begin{pmatrix} 1 & 0 & 0 \\ 0 & -1 & 1 \\ 0 & 0 & 6\end{pmatrix}$$
\end{solution}
\vfill

\newpage
\question The monic Legendre polynomials are an orthogonal family of polynomials on the interval $[-1,1]$.  The first three monic Legendre polynomials are $L_0(x) = 1$, $L_1(x) = x$, and $L_2(x) = x^2 - \tfrac{1}{3}$. The $L^2[-1,1]$ norms of these polynomials are given by $\|L_0\|^2 = 2$, $\|L_1\|^2 = \tfrac{2}{3}$, and $\|L_2\|^2 = \tfrac{8}{45}$.  
\begin{parts}
\part What is the $L^2[-1,1]$ norm of the function $f(x) = x^4$? 
\begin{solution}
$\|f\|^2 = \int_{-1}^1 f(x)^2 \,dx = \frac{2}{9}$, so $\|f\| = \frac{\sqrt{2}}{3}$. 
\end{solution}
\vfill


\part Find the orthogonal projection of the function $f(x) = x^4$ onto the span of the first three monic Legendre polynomials. 
\begin{solution}
\end{solution}
\vfill
\end{parts}

\question The polynomial $p_4(x) = 1 + x + \tfrac{x^2}{2} + \tfrac{x^3}{3!} + \tfrac{x^4}{4!}$ is the 4th degree Maclaurin polynomial for $e^x$.  Suppose you use $p_4(-2)$ to estimate the value of $e^{-2}$.  What is the worst case error in that approximation, according to Taylor's remainder theorem? 
\begin{solution}
The remainder is $\dfrac{e^\xi}{5!} 2^5 \le \dfrac{32}{120}$. 
\end{solution}
\vfill


\question The function $\ds f(x) = e^{-0.1x}$ has a fixed point near $x_0 = 1$.  Suppose you calculate $x_1 = f(x_0)$, then $x_2 = f(x_1)$, and then continue calculating $x_{n+1} = f(x_n)$ recursively. (i) How can you tell that this recursive process will converge? (ii) What is the fixed point, accurate to four decimal places? 
\begin{solution}
(i) It converges because $|f'(x)| = |0.1e^{-0.1x}| \le 0.1$ for all $x >0$.  \\
(ii) Starting with $x_0=1$, I got the fixed point $0.9128$ after just four steps.
\end{solution}
\vfill

\newpage
\question Find two orthogonal vectors that have the same span as $\begin{bmatrix} 1 \\ 3 \\ 0  \end{bmatrix}$ and $\begin{bmatrix} 10 \\ -10 \\ 2  \end{bmatrix}$. 
\vfill

\question Recall the composite trapezoid rule:
$$\int_a^b f(x) \, dx \approx \frac{h}{2}(f(x_0) + 2f(x_1) + 2f(x_2) + \ldots + 2 f(x_{n-1}) + f(x_n) ).$$
If you could calculate the function $f$ exactly every time, then the error in the composite trapezoid rule would be:
$$|\text{Error}| \le \max_{a \le \xi \le b} |f^{(2)}(\xi)| \frac{(b-a)^3}{12n^2}.$$
In reality, you typically cannot calculate the function $f$ exactly because of rounding errors.  
\begin{parts}
\part Assume that every time the function $f$ is computed there is an (absolute) error of up to $\delta \approx 10^{-16}$. What is the worst case total rounding error in the formula above? 
\begin{solution}
Each time you compute $f(x_k)$, you get an additional error of up to $\delta$.  The total after summing and multiplying by $\tfrac{h}{2}$ is 
$$nh\delta = (b-a)\delta,$$ 
in addition to the error formula above. 
\end{solution} 

\vfill
\part What happens to the rounding error in the limit as $n$ gets very large? Does it keep getting smaller or does it start getting bigger eventually? Explain.  
\begin{solution}
Although the theoretical error gets smaller, the error from rounding is constant in the worst case, so it might not go away.  In practice it will probably be a lot smaller than the worst case since the rounding errors will probably tend to cancel out. 
\end{solution}
\end{parts}
\vfill

\question How many floating point operations are needed to multiply two $n$-by-$n$ matrices?  
\vfill

\end{questions}

\end{document}

