\documentclass[10pt]{article}
\usepackage[margin=0.4in]{geometry}
\usepackage{amsmath}
\usepackage{enumitem}
\usepackage{multicol}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric}
\usepackage{soul}

\newcommand{\ds}{\displaystyle}
\newcommand{\on}{\operatorname}


\begin{document}
\newcounter{enumCount}
\pagestyle{empty}
\subsection*{Homework 2 - Math 243 \hfill Name: \underline{\hspace*{2in}}}

\noindent

\begin{enumerate}
\item Suppose that a simple electric circuit has a resistor with resistance $R$ in ohms ($\Omega$), a capacitor with capacitance $C$ in farads (F) and a (time-dependent) voltage source that provides $E(t)$ volts (V).  The voltage drop across the capacitor $E_C$ satisfies the differential equation
$$RC \dfrac{dE_C}{dt} + E_C = E(t).$$
Solve this differential equation if $E(t)$ is a constant $5$ volts, and $R = 2~\Omega$ and $C = 1$ F with initial condition $E_C(0) = 10$ V. 
\vfill

\item Draw a rough sketch of a slope field for the following differential equations. 
\begin{center}
\begin{tikzpicture}
\draw[very thick] (0,4) -- (8,4);
\draw[very thick] (4,8) -- (4,0);
\end{tikzpicture}
\end{center}

\vfill

\item Substitute $e^{at}$ for $y(t)$ in each of the following differential equations.  Then simplify and find all values of the constant $a$ such that $y = e^{at}$ is a solution. 
\begin{enumerate}
\item $y'' + 2y' - 15y = 0$.
\vfill
\item $y''' - 4 y'' + 4 y' = 0$.
\vfill
\end{enumerate}

%\item Recall that $\dfrac{dy}{dt} = - \alpha y$ is the equation for exponential decay.  It can be used to model things like radioactive decay and population decline, among other things.  Consider the following variation of this equation (assume that both $\alpha, \beta >0$):
%$$\dfrac{dy}{dt} = - \alpha y + \beta.$$
%Describe in words one thing that this equation might be useful to model.
%\vfill

\item A stockpile of nuclear waste initially contains 0.8 kilograms of radium. The radium decays exponentially at rate $r$, but new radium waste is added to the stockpile at a rate of 0.02 kilograms per year. Write an differential equation modeling the mass of radium in the stockpile. You don't need to look up the value of the constant $r$. 
\vfill



\newpage
\setcounter{enumCount}{\theenumi}
\end{enumerate}

\noindent
\textit{Solve the following separable equations with the given initial values.}

\begin{multicols}{2}
\begin{enumerate}
\setcounter{enumi}{\theenumCount}
\item $x \, dx - y^2 \, dy = 0$, with $y(0) = 1$. 


\item $\dfrac{dy}{dt} = \dfrac{\cos t}{y}$, with $y(0) = 5$.

\setcounter{enumCount}{\theenumi}
\end{enumerate} 
\end{multicols}

\vfill
\vfill

\begin{multicols}{2}
\begin{enumerate}
\setcounter{enumi}{\theenumCount}
\item $xy' = \sqrt{1-y^2}$, with $y(1) = 0$. \\ \vfill
Hint: recall that $\dfrac{d}{dy} \arcsin(y) = \dfrac{1}{\sqrt{1 - y^2}}$. 



\item $\dfrac{dy}{dt} = y^2$ with $y(0) = 2$.  

\setcounter{enumCount}{\theenumi}
\end{enumerate} 
\end{multicols}

\vfill
\vfill
\vfill





\end{document}
