\documentclass[10pt]{article}
\usepackage[margin=0.4in]{geometry}
\usepackage{amsmath}
\usepackage{enumitem}
\usepackage{multicol}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric}
\usepackage{soul}

\newcommand{\ds}{\displaystyle}
\newcommand{\on}{\operatorname}


\begin{document}
\newcounter{enumCount}
\pagestyle{empty}
\subsection*{Homework 4 - Math 243 \hfill Name: \underline{\hspace*{2in}}}


\noindent
%\textit{Find general solutions for the following differential equations.}
%
%\begin{multicols}{2}
%\begin{enumerate}
%\setcounter{enumi}{\theenumCount}
%\item $\dfrac{dx}{dt} - t^2x = 0$.
%
%
%\item $\dfrac{dy}{dt} - 5y = e^t$.
%
%\setcounter{enumCount}{\theenumi}
%\end{enumerate} 
%\end{multicols}
%\vfill


\noindent
\textit{Solve the following initial value problems using the integrating factors method.}

\begin{multicols}{2}
\begin{enumerate}
\setcounter{enumi}{\theenumCount}
\item $\dfrac{dx}{dt} + 2tx = 4 e^{-t^2}$, $x(0) = 5$.


\item $\dfrac{dy}{dt} + \dfrac{y}{t+1} = 2$, $y(0) = 3$.

\setcounter{enumCount}{\theenumi}
\end{enumerate} 
\end{multicols}
\vfill




\begin{enumerate}
\setcounter{enumi}{\theenumCount}
\item A 30-gallon tank initially contains 10 gallons of water and 4 pounds of dissolved salt.  A concentrated salt water solution containing 1 pound of salt per gallon is added to the tank at a rate of 2 gallons per minute, while a well-mixed solution is drained from the tank at a rate of 1 gallon per minute.  Find a differential equation to model the amount of salt $y(t)$ in the tank.  You do not need to solve the differential equation.
\vfill

\setcounter{enumCount}{\theenumi}
\end{enumerate}

\textit{Find all equilibrium solutions for the following systems of equations.}

\begin{multicols}{2}
\begin{enumerate}
\setcounter{enumi}{\theenumCount}
\item ~ \\ $\begin{array}{c} \dfrac{dx}{dt} = x - 2y + 4 \\ ~ \\ \dfrac{dy}{dt} = 2x + y - 7 \end{array}$.


\item ~ \\ $\begin{array}{c} \dfrac{dR}{dt} = 5R - 0.5 RF \\ ~ \\ \dfrac{dF}{dt} = -4F + 0.1RF \end{array}$.

\setcounter{enumCount}{\theenumi}
\end{enumerate} 
\end{multicols}
\vfill


\newpage
\begin{enumerate}
\setcounter{enumi}{\theenumCount}
\item Consider the differential equation $\dfrac{dv}{dt} + 6v = e^{2t}$.  
\begin{enumerate}
\item What is the corresponding homogeneous differential equation? 
\vfill

\item What is the general solution of the homogeneous differential equation?
\vfill

\item Find one particular solution for the original equation of the form $y(t) = Ae^{2t}$. Then combine the particular solution and the homogeneous solution to express the general solution.  
\vfill
\end{enumerate}





\item Consider the differential equation $\dfrac{dy}{dt} + y = 6 \cos x.$
\begin{enumerate}
\item Find constants $A$ and $B$ such that $y(t) = A \cos t + B \sin t$ is a solution to this ODE.
\vfill


\item Find the general solution of $\dfrac{dy}{dt} + y = 6 \cos x$. 
\vfill
\end{enumerate}

\setcounter{enumCount}{\theenumi}
\end{enumerate}


\end{document}
