\documentclass[10pt]{article}
\usepackage[margin=0.5in]{geometry}
\usepackage{amsmath}
\usepackage{enumitem}
\usepackage{multicol}
\usepackage{tikz}
\usepackage{soul}

\newcommand{\ds}{\displaystyle}

\begin{document}
\newcounter{enumCount}
\pagestyle{empty}
\subsection*{Homework 3 - Math 140 \hfill Name: \underline{\hspace*{2in}}}


\noindent 
\textit{Simplify each expression.}
\begin{multicols}{3}
\begin{enumerate}
\setcounter{enumi}{\theenumCount}
\item $x^2 (5x^7)$.
\item $(4x^3)^2$.
\item $\dfrac{(2x)^3}{6 x^2}$.
\setcounter{enumCount}{\theenumi}
\end{enumerate}
\end{multicols}
\vfill

\noindent
\textit{Simplify and rewrite without negative exponents.}
\begin{multicols}{3}
\begin{enumerate}
\setcounter{enumi}{\theenumCount}
\item $\frac{1}{2} (x^{-3})$.
\item $\left(\dfrac{x^3}{2}\right)^{-3}$.
\item $\dfrac{8 \, x^{-5}}{6 \, x^{-3}}$.
\setcounter{enumCount}{\theenumi}
\end{enumerate}
\end{multicols}
\vfill

\noindent 
\textit{Rewrite using negative and/or fractional exponents, so there are no radical symbols.}
\begin{multicols}{3}
\begin{enumerate}
\setcounter{enumi}{\theenumCount}
\item $\dfrac{3}{\sqrt{x}}$.
\item $x \sqrt[4]{x}$.
\item $\dfrac{x}{\sqrt[3]{x}}$.
\setcounter{enumCount}{\theenumi}
\end{enumerate}
\end{multicols}
\vfill



\newpage

\begin{enumerate}
\setcounter{enumi}{\theenumCount}

\item A clothing business finds there is a linear relationship between the number of shirts, $n$, it can
sell and the price, $p$, it can charge per shirt. In particular, historical data shows that 1000
shirts can be sold at a price of \$30, while 3000 shirts can be sold at a price of \$22 . Find a
linear equation in the form $p = mn + b$ that gives the price $p$ they can charge for $n$ shirts.\vfill

\item How many shirts would the business be able to sell if the price was \$20? 
\vfill

\item If the average cost per shirt is $C = \tfrac{1}{250} n + 6$, then for what production level $n$ is the price per shirt equal to the average cost? 
\vfill
\setcounter{enumCount}{\theenumi}
\end{enumerate}


\noindent 
\textit{Solve the systems of equations below.}
\begin{multicols}{2}
\begin{enumerate}
\setcounter{enumi}{\theenumCount}
\item $\begin{array}{l} 2x - 3y = 1 \\ 3y - x = -2 \end{array}$ 
\item $\begin{array}{l} x + 3y = 0 \\ 2x + 4y = 8\end{array}$ 
\setcounter{enumCount}{\theenumi}
\end{enumerate}
\end{multicols}

\vfill

\end{document}
