---
title: Logistic regression
---

## Example: Predicting success in calculus

The following data includes students' GPAs (both high school and college) and (combined) SAT scores with whether or not they passed Math 140 at Hampden-Sydney.

```{r}
calc <- read.csv("https://people.hsc.edu/faculty-staff/blins/StatsExamples/calculusData.csv")
head(calc)
```

In this case the two explanatory variables (SAT score and high school GPA) are both quantitative, but the response variable is a binary categorical variable which is recorded as 1 for a passing grade, and 0 for a failing grade). 

### One variable model

We'll start with a logistic regression model using just high school GPA as a predictor variable.

```{r}
model1 <- glm(pass ~ hs_gpa, data = calc, family = "binomial")
model1
```

Here is a plot of the data and the logistic regression model.

```{r}
plot(calc$hs_gpa, calc$pass, xlab = "High school GPA", ylab = "Probability of passing calculus", pch = 16)
curve(predict(model1, data.frame(hs_gpa = x), type='response'), add=T)
```


You can use this logistic regression model to predict the log-odds of a student passing Math 140.

```{r}
predict(model1, data.frame(hs_gpa = 2.0))
```

You still have to use the formulas 
$$\text{odds} = \exp(\text{log-odds}) ~~~~\text{ and } ~~~~ p = \frac{\text{odds}}{\text{odds} + 1}$$
to convert the log-odds into the probability that a student passes. 

```{r}
log.odds <- predict(model1, data.frame(hs_gpa = 2.0))
exp(log.odds)/(exp(log.odds) + 1)
```


### Two variable model

```{r}
model2 <- glm(pass ~ hs_gpa + sat, data = calc, family = "binomial")
model2$coefficients
exp(model2$coefficients)

```

This model has the form
$$\log \left( \text{odds of passing} \right) = `r model2$coefficients["(Intercept)"]` + `r model2$coefficients["hs_gpa"]` \text{hs_gpa} + `r model2$coefficients["sat"]` \text{sat}.$$ 

Once again, we can use it to make predictions about a single student. 

```{r}
predict(model2, data.frame(hs_gpa = 2.0, sat = 1000))
```

<br>
<br>

