---
title: Leukemia data
---

Consider the following data published on n = 27 leukemia patients. The response variable `REMISS` is 1 if leukemia remission occurred, 0 otherwise.

The predictor variables are cellularity of the marrow clot section (`CELL`), smear differential percentage of blasts (SMEAR), percentage of absolute marrow leukemia cell infiltrate (`INFIL`), percentage labeling index of the bone marrow leukemia cells (`LI`), absolute number of blasts in the peripheral blood (`BLAST`), and the highest temperature prior to start of treatment (`TEMP`).

```{r}
leukemia <- read.csv("leukemia.csv")
mymodel <- glm(REMISS ~ CELL + INFIL + LI + BLAST + TEMP, data = leukemia, family = "binomial")
summary(mymodel)
```

```
## glm(formula = REMISS ~ CELL + INFIL + LI + BLAST + TEMP, family = "binomial", 
##     data = leukemia)
## 
## Coefficients:
##             Estimate Std. Error z value Pr(>|z|)  
## (Intercept)   77.523     69.650   1.113   0.2657  
## CELL          10.582      9.335   1.134   0.2570  
## INFIL          1.246      3.878   0.321   0.7480  
## LI             4.286      2.517   1.703   0.0886 .
## BLAST         -0.227      2.201  -0.103   0.9178  
## TEMP         -94.013     74.867  -1.256   0.2092  
```

```{r}
mymodel2 <- glm(REMISS ~ LI, data = leukemia, family = "binomial")
summary(mymodel2) 
```

```{r}
x <- log(rbeta(100, 1, 1))
qqnorm(x)
qqline(x)
```
