---
title: "Multiple Regression: Model Selection"
output: slidy_presentation
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

## Introduction

When we choose a model, we have two goals: 

* **Predictive power** How precise and accurate are the predictions?
* **Parsimony** How simple is the model?

These seem like they would be conflicting goals, but sometimes simple models can have a lot predictive power.

## Reasons to remove variables

To make our models parsimonious, we remove variables. Here are reasons to remove variables:

* **Overfitting** - including lots of variables can lead to **overfitting**, i.e., modeling the noise from the sample, not the underlying trend. 


* **Insignificant variables** - some variables might not have a statistically significant effect on the response, so it usually doesn't make sense to include them in the model.

* **Multicolinearity** - sometimes two input variables are highly correlated, and including them both in the model is redundant.  

## Measuring predictive power

In multilinear regression, we often have variables that aren't statistically significant contributors to the model. We typically want to remove these variables. 

Recall that 
$$R^2 = \frac{SSM}{SST} \text{ and } 1 - R^2 = \frac{SSE}{SST}.$$

One way to measure the predictive power of a model is to use the **adjusted $R^2$** which is defined by

$$R^2_\text{adj} = 1 - \frac{MSE}{MST}.$$



## Example: Babies birth weights

```{r}
babies <- read.csv('http://people.hsc.edu/faculty-staff/blins/classes/spring17/math222/data/babies.csv')
dim(babies)
```

The `babies` data frame contains data on baby birth weights. The variables are:

* bwt - birth weight (in ounces)
* gestation - length of the pregnancy (in days)
* parity - 1 if baby was first born, 0 otherwise
* age - mother's age (in years)
* height - mother's height (in inches)
* weight - mother's weight (in lbs.)
* smoke - 1 if the mother is a smoker, 0 otherwise


## The linear model

Recall from last time that 62 rows out of 1236 were incomplete, so we won't include those in the linear model. 

```{r}
babies_lm <- babies |>
    na.omit() |>
    lm(bwt ~ gestation + parity + age + height + weight + smoke, data = _)  
babies_lm
```


<br>
<br>

## Which variables to include?

Ideally, you would every subset of variables and see which subset produces the largest $R^2_\text{adj}$.  That is time consuming, even on a computer, when there are a large number of variables.  So our textbook suggests two alternatives:

* **Forward selection** - add variables to the model one-at-a-time, always picking the variable that increases $R^2_\text{adj}$ the most.  Stop when you cannot increase $R^2_\text{adj}$ any more. 
* **Backward selection** - start with the full model, and remove variables one-at-a-itme, always removing the variable that increases $R^2_\text{adj}$ the most.  Stop when you cannot increase $R^2_\text{adj}$ any more. 

<br>
<br>


## Practice 

```{r}
babies <- read.csv('http://people.hsc.edu/faculty-staff/blins/classes/spring17/math222/data/babies.csv') |>
    na.omit()
babies_lm <- lm(bwt ~ gestation + parity + age + height + weight + smoke, data = babies) 
adj.R.sq0 = summary(babies_lm)$adj.r.squared
```

So the initial $R^2_\text{adj} = `r adj.R.sq0`$.

1. Starting with the code above, which variable (if any) should you remove first to increase $R^2_\text{adj}$ the most? 

```{r}
new_lm <- lm(bwt ~ gestation + parity + height + weight + smoke, data = babies) 
summary(new_lm)$adj.r.squared
```

## Alternative approach

If all we care about is predictive power, then maximizing $R^2_\text{adj}$ is a good strategy.  But if we want to better understand which variables really matter, then we might want to focus on the p-values for each variable in the model instead. 

```{r}
summary(new_lm)
```

<br>
<br>

## Another Example

Our textbook has a case study about Ebay auctions of Mario Kart for the Nintendo Wii. The outcome
variable of interest is the total price (`price`) of an auction, which is the highest bid plus the shipping
cost. Among the other variables are the following:

* `cond_new`: Indicator variable for if the game is new (1) or used (0).
* `stock_photo`: Indicator variable for if the auction’s main photo is a stock photo.
* `duration`: The length of the auction, in days, taking values from 1 to 10.
* `wheels`: The number of Wii wheels included with the auction. A Wii wheel
is an optional steering wheel accessory that holds the Wii controller.


## Forward selection

2. Use forward selection to decide which of the variables `cond_new`, `stock_photo`, `duration`, or `wheels` to add to the model first. Then keep adding variables until you can't increase $R^2_\text{adj}$ any more. 

```{r}
mario <- read.csv("mariokart.csv")
mario_lm <- lm(price ~ cond_new + stock_photo + wheels, data = mario)
summary(mario_lm)$adj.r.squared
```

3. How much would you predict for the total price for the Mario Kart game if it was used, used a stock
photo, and included two wheels and put up for auction during the time period that the Mario Kart
data were collected?  Make a 90% prediction interval.

4. Make a plot of the `mario` data frame and use it to check to see if the linearity assumptions appears valid.

5. Make a plot of the `mario_lm` and use it to assess whether the constant variance and normal residuals assumptions hold. 

## Checking linearity

We need to check that there is a roughly linear relationship between each of the explanatory variables and the response variable. This also lets you see how the residuals depend on each explanatory variable.


```{r}
plot(mario)
```

<br>
<br>

## Checking variance & residuals

```{r}
plot(mario_lm, which = 1:2)
```



