\documentclass[12pt]{article}

\usepackage[margin=0.5in]{geometry}
\usepackage{amsmath}
\usepackage{tikz}
\usepackage{hyperref}

\newcommand{\ds}{\displaystyle}

\begin{document}
\pagestyle{empty}
\subsubsection*{Midterm 2 Review Problems \hfill Math 141 }
\textit{These are suggested review problems similar to what might be on Midterm 2. Included with each problem is a link to a video where you can see how the problem is solved. I didn't make the videos, they are all  available on YouTube.}

\begin{enumerate}
\item Find the derivative of $f(x) = \dfrac{36x^5 - 12 x^4 + 6}{3x^2}$. 
\vfill 
\hfill  \url{https://youtu.be/PZXNRzLgXXU} 
\hrule 

\item Find the derivative of $f(x) = (x^3 + 2x^2 - x^{-2})^{-7}$.
\vfill 
\hfill  \url{https://youtu.be/6_lmiPDedsY} 
\hrule 

\item Find the derivative of $y = \sec \theta  \tan \theta$.
\vfill
\hfill \url{https://youtu.be/uNk7NoVlWlQ}
\hrule


\item Use implicit differentiation to find the derivative $y'$ when $ 4 \left( \cos x \sin y \right) = 1$.
\vfill
\hfill \url{https://youtu.be/fUNfxxOI1Rs}
\hrule

\newpage
\item Suppose that $C(A)$ is the cost of building a house as a function of its area in square feet.  For each of the following equations, give a simple one sentence explanation (including units) of what exactly the equation is saying. 
\begin{enumerate}
\item $C(1{\small,}000) = 300{\small,}000$
\item $C'(1{\small,}000) = 350$ 
\end{enumerate}
\vfill
\hfill \url{https://youtu.be/QirtTPD0Unk?t=138}
\hrule

\item Find the first, second, and third derivatives of the function $y = \dfrac{6}{(x-1)^2}$. 
\vfill
\hfill \url{https://youtu.be/yIfA_ux1cos}
\hrule


\item Use a linear approximation to estimate $\sqrt[3]{1001}$.  Hint: use the linearization of $f(x) = \sqrt[3]{x}$ with $a = 1000$. 
\vfill
\hfill \url{https://youtu.be/KO4mGU5_ZkU}
\hrule

\item The radius of a sphere is measured to be $r = 5$ meters with an error of $dr = 0.05$ meters. Find the differential of the volume $V = \tfrac{4}{3} \pi r^3$ and use it to estimate the error in the computed volume. 
\vfill
\hfill \url{https://youtu.be/O05Rg4cvXWM}
\hrule


\newpage
\item A cone is being filled with 1 cubic centimeter of water per second.  If the cone is 4 cm tall and has a diameter of 4 centimeters, then how fast is the water level rising when the water level is 2 cm?  Hint: The volume of the water in the cone is $V = \tfrac{1}{12} \pi h^3$ where $h$ is the water level in centimeters.  
\vfill 
\hfill  \url{https://youtu.be/Xe6YlrCgkIo}
\hrule 


\item Find the absolute max and min for $f(x) = x + \dfrac{1}{x}$ on the interval $[\tfrac{1}{5}, 4].$
\vfill 
\hfill  \url{https://youtu.be/enaSoNJTnFM} 
\hrule 


\item Find the critical points of $f(\theta) = 4 \sec \theta + 2 \tan \theta$ on the interval $0  < \theta < 2\pi$. 
\vfill 
\hfill  \url{https://youtu.be/MRG1X4ld2As?t=87} 
\hrule 


\item A hammer is thrown upwards from the surface of the moon.  Its height is given by $s(t)  = 32t - 0.8t^2$ where $s$ is in meters and $t$ is in seconds.  How long until the hammer stops going up and starts falling back down? And what is the maximum height of the hammer?
\vfill 
\hfill  \url{https://youtu.be/R1H0i5Dx8IE} 
\hrule 

\end{enumerate}


\end{document}














