\documentclass[11pt,answer]{exam}
\usepackage[margin=0.7in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{colortbl}
\usepackage{multicol}
\usepackage{tikz,pgfplots}
\usepackage{soul}
\usepackage{circledsteps}

\newcommand{\blank}[1]{\underline{\hspace*{#1}}}
\newcommand{\ds}{\displaystyle}
\newcommand{\mymod}{~\mathrm{mod}~}


\begin{document}
\pagestyle{empty}
\graphicspath{{/home/brian/Dropbox/HSC/Spring16/Math111/}}

\subsection*{Midterm 2 Review \hfill Math \& Society}

\textit{The following problems are similar to ones you might see on the midterm exam. There is a also a list of terms \& facts you should memorize on the last page.}

\begin{questions}

\question \hl{Apportionment question - Hamilton's method}

\question \hl{Find an Euler path if there is one...}

\question Draw a tree that has a degree 4 vertex, a degree 3 vertex, and a degree 2 vertex. 
\begin{solution}
Test
\end{solution}



\question Use the mileage chart below to find the minimum spanning tree connecting these five cities.  
\begin{center}
\begin{tabular}{l|ccccc}
& Boston & Buffalo & Chicago & Columbus & Louisville \\ \hline
Boston & -- & 446 & 963 & 735 & 941 \\ 
Buffalo & 446 & -- & 522 & 326 & 532 \\ 
Chicago & 963 & 522 & -- & 308 & 292 \\
Columbus & 735 & 326 & 308 & -- & 209 \\
Louisville & 941 & 532 & 292 & 209 & -- 
\end{tabular}
\end{center}
\begin{solution}

\end{solution}

\question Which of the following graphs must be a tree?  Circle all that are.  
\begin{choices}
\choice An ancestry graph.  People are vertices; edges connect parents to their children.
\choice Friends network.  People are vertices; edges connect people who are friends.  
\choice Railway network. Cities are vertices; railway routes are the edges.
\choice Markov chain graph strongly connected components.  Strongly connected components are the vertices; Two components share an edge if it is possible to get from one to the other.  
\end{choices}


\end{question}

\end{document}
