\documentclass[11pt]{exam}
\usepackage[empty]{fullpage}
\usepackage{amsmath,amssymb}
\usepackage{multicol}
\usepackage{tikz}
\usepackage{soul}

\newcommand{\ds}{\displaystyle}

\newcounter{countA}

\begin{document}
\subsubsection*{Midterm 3 Review - Math 243}

\begin{questions}
\question Consider the system
$$\frac{dx}{dt} = x^2 - y$$
$$\frac{dy}{dt} = x^2 - 1$$
\begin{parts}
\item Find all equilibria.  
\vfill

\item Use linearization (i.e., the Jacobian matrix) to classify each equilibrium point.  
\vfill

%\item Which of the following phase portraits might represent the system above?  
%\vfill
\end{parts}
\newpage
%%%%%%%%%%%%%%%%%
%%% Problem 2 %%%
%%%%%%%%%%%%%%%%%
\question Consider the following one-parameter family of systems.  
$$\frac{dx}{dt} = x^2 - y$$
$$\frac{dy}{dt} = x^2 - a$$
\begin{parts}
\item Find all equilibrium points.
\vfill
\item Identify all values of $a$ at which a bifurcation occurs.   
\vfill
\item Briefly describe how the equilibrium points change at each of the bifurcation values.  
\vfill

\end{parts}
\newpage
%%%%%%%%%%%%%%%%%
%%% Problem 3 %%%
%%%%%%%%%%%%%%%%%
\question Consider the system 
$$\dfrac{dx}{dt} = x \cos(xy)$$
$$\dfrac{dy}{dt} = -y \cos(xy)$$
\begin{parts}
\item Show that the quantity $H(x,y) = \sin(xy)$ is a Hamiltonian function for the system.
\vfill

\item Sketch the level sets for this system.  Hint: The level sets occur when $xy$ is a constant.  What does the graph of $xy = c$ look like?  
\vfill

\item Sketch the solution curve that passes through the point $(1,\pi)$.  Be sure to indicate the direction of the solution.
\vfill

\end{parts}
\newpage
%%%%%%%%%%%%%%%%%
%%% Problem 4 %%%
%%%%%%%%%%%%%%%%%
\question Consider the following forced harmonic oscillator.  
  $$y'' + 3y' + 2y = \cos 2t$$
\begin{parts}
\item What is the complexification of this differential equation?  
\vfill

\item Find a particular solution for this differential equation.  
\vfill

\item Find the general solution for this differential equation.
\vfill

\item If this system represents a harmonic oscillator, is it overdamped or underdamped?  How can you tell?  
\vfill
\end{parts}


\newpage
%%%%%%%%%%%%%%%%%
%%% Problem 5 %%%
%%%%%%%%%%%%%%%%%
\question Consider the non-homogeneous linear system
$$\frac{d \mathbf{y}}{dt} =   A \mathbf{y}  + \mathbf{b} , \text{ where } A = \begin{bmatrix} 1 & 3 \\ -1 & 1 \end{bmatrix}\text{ and } \mathbf{b} = \begin{bmatrix} -6 \\ 2 \end{bmatrix}.$$
\begin{parts}
\item Find the equilibrium point for the system.  
\vfill

\item Find the general solution to the system above.  
\vfill

\item If you replace the constant vector $\mathbf{b}$ above with the vector-valued function $\mathbf{b}(t) = \begin{bmatrix} -6e^{-t} \\ 2 e^{-t} \end{bmatrix}$. What would make a good guess for a particular solution $\mathbf{y}_p$ for this new system? 
\vfill
\end{parts}

\question Find general solutions to the following non-homogeneous equations. 

\begin{multicols}{2}
\begin{parts}
\part $y'' - y' - 6 = e^{4t}$


\part $y'' - y' - 6 = e^{2t}$

\setcounter{countA}{\value{partno}}
\end{parts}
\end{multicols}
\vfill

\begin{multicols}{2}
\begin{parts}
\setcounter{partno}{\value{countA}}
\part $y'' + 4y = \cos(2t)$
\vfill

\part $y'' + 2y' + 6y = 3t+2$
\vfill

\end{parts}
\end{multicols}



\end{questions}

\end{document}
